---
title: "Cluster Results"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
runtime: shiny
---

```{r, warnings=FALSE, echo=FALSE, messages=FALSE}
source('../util.R')
library(ggplot2)
library(knitr)
library(shiny)
```

```{r}
sample <- 1000
seed <- 42
dir <- file.path('../data', paste(sample, seed, sep='_'))
comments <- getClusterResults(dir)
```

```{r fig1, fig.width=12, fig.height=8}
ggplot(comments, aes(cluster_id)) +
    facet_wrap(~method, scales='free_x', ncol=4) + 
    geom_bar() +
    ggtitle('Count of clusters by method')
```

```{r table_1}
cluster_fns <- getClusters(dir)
clusters <- lapply(cluster_fns, function(fn) {
    key <- sub('clusters_(.*).csv', '\\1', fn)
    read.csv(file.path(dir, fn)) %>% 
        mutate(method=key) %>%
        replace_na(list(membership_prob=0)) %>%
        group_by(method) %>%
        summarize(clusters_count=length(unique(cluster)),
                  mean_prob=round(mean(membership_prob), 2),
                  median=round(median(membership_prob), 2),
                  max=round(max(membership_prob), 2))
}) %>% bind_rows()

cluster.table <- tableGrob(clusters, rows = NULL)
grid.draw(cluster.table)
```

When comparing the membership probabilities by method, we see the methods that produced the most clusters were on the lower side of the mean probability, `mean_prob`. Actually, when dealing with ngrams, HDBSCAN membership probability for a single point never peaked above 85%. For the most part, these methods produced fewer clusters than the groups identified by Kao; although, it was never explicitly mentioned in his analysis that one cluster from a signature string match equated to one unique author.

```{r cluster_explorer}
shinyAppDir(
    '.',
    options = list(width = '100%', height = 1000)
)
```